### 2021-11-25 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@ef1c555be2d8ab4689ed048597fb1e7a05131721 🚀

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
observing an interaction between species in order to aid in the design
of samples of ecological interactions across space, and to provide a
null expectation of the false-negative probability for a dataset of a
given size. Further, it then proposes a method for optimizing the
spatial sampling locations to maximize the probability of detecting an
interaction between two species given a fixed number of total of
observations and the distributions of each species. This addresses the
optimization of monitoring network part of the flow from data to
mitigation at the top of fig.~\ref{fig:thesis}, left. There is more than
one way to observe a false-negative when sampling interactions---
fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of false-negatives in
occurrence, co-occurrence, and interaction data. The goal of this
chapter is to untangle the relationship between co-occurrence
false-negatives (and their relation to relative abundance) and
interaction false-negatives.

@@

@@

Interactions between plants and pollinators form networks which together
structure the ``architecture of biodiversity'' (Bascompte \& Jordano
2007). The functioning and stability of ecosystems emerge from these
interactions, but anthropogenic change threatens to unravel and
``rewire'' these interaction networks (CaraDonna \emph{et al.} 2017),
jeopardizing the persistence of these systems. Plant-pollinator networks
face two possible forms of rewiring in response to anthropogenic
environmental change: spatial and temporal. Range shifts could cause
interacting species to no longer overlap in space, and shifts in
phenology could cause interacting species to no longer occur at the same
time of year. This chapter uses several years of data on
bumblebee-flower phenology and interactions across several field sites,
each consisting of several plots across an elevational gradient,
combined with spatial records of species occurrence via GBIF to forecast
the uncoupling of the plant-pollinator metaweb of Colorado.

@@

@@

As land-use change has caused many habitats to become fragmented and
patchy, promoting landscape connectivity has become of significant
interest to mitigate the effects of this change on Earth's biodiversity.
However, the practical realities of conservation mean that there is a
limitation on how much we can modify landscapes in order to do this. So
what is the best place to put a corridor given a constraint on how much
surface-area you can change in a landscape? This is the question this
chapter seeks to answer. Models for inferring corridor locations have
been developed, but are limited in that are not developed around
promoting some element of ecosystem function, but instead by trying to
find the path of least resistance in an existing landscape from a
derived resistance surface (e.g. Peterman 2018). This chapter proposes a
general algorithm for optimizing spatial restoration effort to move a
measurement of ecosystem functioning toward a target state.

@@

@@

The final chapter consists of a collection of modules in the
\texttt{Julia} language for different aspects of community ecology,
including most of the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is at the center of this set
of tools, but due to the nature of the Julia language, MCD.jl is
inter-operable with several existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. We need a software library like this to generate synthetic
data from a \emph{known} set of mechanisms and parameters to test our
methods for parameter inference and forecasting on this \emph{known}
system to assess the effectiveness of these inference and forecasting
methods.

@@

@@

@@

>>>>

\hypertarget{context}{%
\subsection{Context}\label{context}}

Ecosystems are composed of interactions between species, yet this
remains one of the greatest shortfalls of data in biodiversity science
(Strydom \emph{et al.} 2021b). This is, in no small part, because
species interactions are considerably more difficult to sample than
single-species processes. However, understanding and predicting species
interactions, and how anthropogenic change is affecting them, is one of
the grand challenges faced by modern ecologists (Jordano 2016). Further,
co-occurrence is often used as a proxy for species interaction when this
is not necessary always a valid assumption (Blanchet \emph{et al.}
2020).

@@

The goal of this chapter is to use simulation models to investigate the
relationship between species relative abundance, sampling effort, and
probability of observing an interaction between species in order to aid
in the design of samples of ecological interactions across space, and to
provide a null expectation of the false-negative probability for a
dataset of a given size. Further, it then proposes a method for
optimizing the spatial sampling locations to maximize the probability of
detecting an interaction between two species given a fixed number of
total of observations and the distributions of each species. This
addresses the optimization of monitoring network part of the flow from
data to mitigation at the top of fig.~\ref{fig:thesis}, left.

@@

\hypertarget{context-1}{%
\subsection{Context}\label{context-1}}

Interactions between plants and pollinators form networks which together
structure the ``architecture of biodiversity'' (Bascompte \& Jordano
2007). The functioning and stability of ecosystems emerge from these
interactions, but anthropogenic change threatens to unravel and
``rewire'' these interaction networks (CaraDonna \emph{et al.} 2017),
jeopardizing the persistence of these systems. Plant-pollinator networks
face two possible forms of rewiring in response to anthropogenic
environmental change: spatial and temporal. Range shifts could cause
interacting species to no longer overlap in space, and shifts in
phenology could cause interacting species to no longer occur at the same
time of year.

@@

This chapter uses several years of data on bumblebee-flower phenology
and interactions across several field sites, each consisting of several
plots across an elevational gradient, combined with spatial records of
species occurrence via GBIF to forecast the uncoupling of the
plant-pollinator metaweb of Colorado.

@@

\hypertarget{context-2}{%
\subsection{Context}\label{context-2}}

As land-use change has caused many habitats to become fragmented and
patchy, promoting landscape connectivity has become of significant
interest to mitigate the effects of this change on Earth's biodiversity
(Resasco 2019). However, the practical realities of conservation mean
that there is a limitation on how much we can modify landscapes in order
to do this.

@@

What is the best place to put a corridor given a constraint on how much
surface-area you can change in a landscape? This is the question this
chapter seeks to answer. Models for inferring corridor locations have
been developed, but are limited in that are not developed around
promoting some element of ecosystem function, but instead by trying to
find the path of least resistance in an existing landscape from a
derived resistance surface (e.g. Peterman 2018). This chapter proposes a
general algorithm for optimizing spatial restoration effort to move a
measurement of ecological dynamics toward a target state.

@@

\hypertarget{context-3}{%
\subsection{Context}\label{context-3}}

Building models for ecological forecasting is an imperative (Dietze
\emph{et al.} 2018), but before applying forecasting methods to actual
data, we need a way to determine the efficacy of different forecasting
methods. We need a software library like this to generate synthetic data
from a \emph{known} set of mechanisms and parameters to test our methods
for parameter inference and forecasting on a system with \emph{known}
ground-truth dynamics to assess the effectiveness of these inference and
forecasting methods.

@@

The final chapter consists of a collection of modules in the
\texttt{Julia} language for different aspects of community ecology,
including most of the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is at the center of this set
of tools, but due to the nature of the Julia language, MCD.jl is
inter-operable with several existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed.

@@

\leavevmode\hypertarget{ref-Dietze2018IteNea}{}%
Dietze, M.C., Fox, A., Beck-Johnson, L.M., Betancourt, J.L., Hooten,
M.B., Jarnevich, C.S., \emph{et al.} (2018). Iterative near-term
ecological forecasting: Needs, opportunities, and challenges.
\emph{Proceedings of the National Academy of Sciences}, 115, 1424--1432.

@@

\leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%
Jordano, P. (2016). Chasing Ecological Interactions. \emph{PLOS
Biology}, 14, e1002559.

@@

\leavevmode\hypertarget{ref-Resasco2019MetDec}{}%
Resasco, J. (2019). Meta-analysis on a Decade of Testing Corridor
Efficacy: What New Have we Learned? \emph{Current Landscape Ecology
Reports}, 4, 61--69.
